\sysfull

### 2017-11-14 paper/coalescing.tex

\textbf{Task Coalescing Strategies.} The simplest strategy will coalesce/de-coalesce fixed number of task at every successful/unsuccessful commit operation, where task coalescing factor increases by $C_U$ at each successful virtual task execution and decreases by $C_D$, otherwise. There are naturally infinite possible strategies for coalescing, i.e. infinite amount of $(C_U,C_D)$ pairs. The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions. A \emph{good} coalescing strategy must moderates the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work. Therefore we propose to consider two fundamental coalescing strategies for \sys:

\begin{enumerate}
	\item \textbf{Power Interrupt History-driven Coalescing:} ; and 
	\item \textbf{Power Interrupt History-driven Coalescing with Weights:} where coalescing strategy takes into account \emph{expected size} of each task in the application to pick the best coalescing factor.
\end{enumerate}

\subsection{Operation of Power Interrupt History-driven Coalescing.} \sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the current operational period. \sys begins execution in a coalesced task with $C$ initialized to half of the number of tasks executed during the previous operating period and $S$ initialized to $0$. The system executes $C-1$ coalesced tasks followed by one committing task. After each commit, \sys assigns $C$ to half of $S$. If more tasks complete in an operating period, \sys attempts to coalesce more tasks. If fewer tasks complete, \sys attempts to coalesce fewer tasks. If no committing task completes due to a power failure, \sys updates $C$ on reboot to half of $S$ (the number of tasks completed before the power failure) and continues.

\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys. At left, there is no coalescing and tasks incur the cost of commit at each transition. The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit.  At right, \sys
coalesces three tasks ($C=3$) avoiding two commits. Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value
of $C$ based on $S$. \sys executes tasks in the first operating period, with
$C$ equal to half the number of tasks completed in the prior execution period.
In the example, $C$ starts at eight. After completing eight tasks, \sys commits
and reduces $C$ by half to four.  \sys completes two of four tasks and power
fails.  During the operating period, \sys executed a total of ten tasks (i.e.,
$S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$).
\sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$
to half of $S$, which is three (after rounding up). \sys coalesces three tasks
and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution
continues.

\subsection{Operation of Program Profiling-driven Coalescing:} \todo{To be written}{Przemek}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}
	\caption{Result of benchmark applications (Section~\ref{}) task size measurement, measured in the number of instruction cycles. We observe very large variations of task sizes per each benchmark. \todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
\label{fig:task_profiling}
\end{figure}

Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}. Each benchmark was run for predefined period of time, on non-intermittent power, and at each task evocation, its execution time was measured. If a task was evoked more than once, longer execution time replaces the short one---this way we obtain conservative values of task sizes. At the end of task measurement, normalized values of task execution are assigned for each task, the shortest one being one task, the longest one X tasks.

\begin{figure}
	\centering
	%\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
	%\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
	\caption{\sys coalesces tasks to avoid overheads.}
	\label{fig:coalescing_example}
\end{figure}

>>>>

\textbf{Task Coalescing Strategies.} The simplest strategy will coalesce/de-coalesce fixed number of task at every successful/unsuccessful commit operation, where task coalescing factor increases by $C_U$ at each successful virtual task execution and decreases by $C_D$, otherwise. There are naturally infinite possible strategies for coalescing, i.e. infinite amount of $(C_U,C_D)$ pairs. The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions. A \emph{good} coalescing strategy must moderates the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work. Therefore we propose to consider two fundamental coalescing strategies for \sys, which we will later assess and compare experimentally:

\begin{enumerate}
	\item \textbf{Power Interrupt History-based Coalescing (HDC):} when more tasks are coalesced without power interrupt $C_U$ increases gradually, and $C_D$ increases, as more task are interrupted before committing; and 
	\item \textbf{Weighted Power Interrupt History-based Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
\end{enumerate}

\subsection{Power Interrupt History-driven Coalescing}

With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period. \sys begins execution in a coalesced task with $S_t=0$. When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows. After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$. This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks. However, tasks coalescing are increased by a half the speed of the successful task completion. This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt. In the power failure case, if fewer tasks complete, \sys attempts to coalesce fewer tasks.

%\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys. At left, there is no coalescing and tasks incur the cost of commit at each transition. The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit. At right, \sys coalesces three tasks ($C=3$) avoiding two commits. Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$. \sys executes tasks in the first operating period, with $C$ equal to half the number of tasks completed in the prior execution period. In the example, $C$ starts at eight. After completing eight tasks, \sys commits and reduces $C$ by half to four. \sys completes two of four tasks and power fails.  During the operating period, \sys executed a total of ten tasks (i.e., $S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$). \sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$ to half of $S$, which is three (after rounding up). \sys coalesces three tasks and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution continues.

\subsection{Weighted Power Interrupt History-driven Coalescing}

The previous coalescing strategy is oblivious to the task size. Although many tasks can be coalesced, their non-equal size will determine that some coalesced tasks will be more susceptible to power interrupts than others. To understand that effect we provide the following illustration.

\textbf{Task Size Profiling.} To measure how large the distribution of tasks is, we have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginnin of Section~\ref{sec:methodology_evaluation} for system setup). Each benchmark was run for predefined period of time (X\,seconds \todo{Provide exact number}{Amjad}), on non-intermittent power, and at each task evocation, its execution time was measured. If a task was evoked more than once, longer execution time replaced the short one---this way we obtained conservative values of task sizes. At the end of task measurement, normalized values of task execution are assigned for each task, the shortest one being one task, the longest one X tasks \todo{Provide exact number}{Amjad}. Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}. We observe huge variations of task sizes---both within a benchmark, as well per benchmark, which calls for the use of this information in coalescing.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}%coalescing_timeline
	\caption{Result of benchmark applications (Section~\ref{sec:results_software}) task size measurement, measured in the number of instruction cycles. We observe very large variations of task sizes per each benchmark. \todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
	\label{fig:task_profiling}
\end{figure}

In the second coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device.

%\begin{figure}
%	\centering
	%\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
	%\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
	%\caption{\sys coalesces tasks to avoid overheads.}
	%\label{fig:coalescing_example}
%\end{figure}
